Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-15 nips_2018.tex
put list functions first

We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
\begin{equation}
    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\hspace{1cm}  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)
  \end{equation}
where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
$\lambda$ acts as a regularizer on the size of the DSL,
and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this  may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
our DSLs are context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting it all together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(fold a b (lambda (x z) }\\
      \phantom{\code{(fold a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}

@@

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example, we apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

\section{Sequence manipulating programs}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

@@

\begin{table}
\centering
\begin{tabular}{lll}
  \toprule
  Name & Input & Output \\\midrule
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
  \bottomrule
\end{tabular}
\caption{Sample tasks from our list function domain %% We used 15 input/output
  %% examples for each task to reduce ambiguity in the task's function,
  %% ensuring solutions achieve the desired concept.
}
\label{listexamples}
\end{table}

>>>>

To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this  may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
our DSLs are context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(fold a b (lambda (x z) }\\
      \phantom{\code{(fold a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\begin{comment}
\begin{table}
\centering
\begin{tabular}{lll}
  \toprule
  Name & Input & Output \\\midrule
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
  \bottomrule
\end{tabular}
\caption{Sample tasks from our list function domain %% We used 15 input/output
  %% examples for each task to reduce ambiguity in the task's function,
  %% ensuring solutions achieve the desired concept.
}
\label{listexamples}
\end{table}

  \end{comment}

@@

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}
